---
title: "Homework 1"
author: "PLSC 31101"
date: "Fall 2019"
output: html_document
---

## Assignment 2

- __Assigned__:
- __Due__: 

For this assignment, you'll use what you know 

### 1. Basics {-}

```{r}
setwd(dir="~/Dropbox/berkeley/Git-Repos/PS239T/06_r-basics/")
```

#### 1.2 Remove all objects from your environment. {-}

```{r}
rm(list=ls())
```

### 2. Vectors and Lists {-}

#### 2.1 Create three vectors: {-}

* a character vector, `titles`, that contain the names of 3 of your favorite movies

* a numeric vector, `year`, that contains the years in which those movies were produced

* a boolean vector `bechdel` that TRUE/FALSE according to whether those movies pass the [bechdel test](http://bechdeltest.com/)

```{r}
titles <- c("Dog Day Afternoon", "The Graduate", "Breakfast Club")
year <- c(1975, 1967, 1985)
bechdel <- c(TRUE, FALSE, TRUE)
```

#### 2.2 Put those three vectors in a list, called `movies`. {-}

```{r}
movies <- list(titles, year, bechdel)
```

#### 2.3 Print the structure of the list `movies` {-}

```{r}
str(movies)
```

### 3. Factors {-}

Here's some code that prints a simple barplot:

```{r}
f <- factor(c("low","medium","high","medium","high","medium"))
table(f)
barplot(table(f))
```

#### 3.1 How would you relevel `f` to be in the correct order? {-}

```{r}
f <- factor(f, levels = c("low", "medium", "high"))

# Test your code
barplot(table(f))
```

### 4. Dataframes {-}

#### 4.1 Coerce the `movies` list you made above into a dataframe. Call it `movies_df` {-}

```{r}
movies_df <- as.data.frame(movies)
```

#### 4.2 Add appropriate column names to `movies_df` {-}

```{r}
names(movies) <- c("film", "year", "bechtel")
```