---
title: "Homework 1"
author: "PLSC 31101"
date: "Fall 2019"
output: html_document
---

## Assignment 2

- __Assigned__:
- __Due__: 

For this assignment, you'll use what you know about R syntax and data structures to perform some common data operations.

### 1. Basics {-}

#### 1.1 Fix the following syntax errors. Enter your corrected code in the second chunk. {-}

```{r eval = F}
# 1
states <- ("California", "Illinois", "Ohio")

# 2
countries <- c("Iran", "Indonesia," "India", "Italy")

# 3
df <- data.frame(age = c(21, 66, 35)
                 party = c('rep', 'dem', 'rep'))

# 4
my-vector <- c("apples", "oranges", "kiwis")

# 5
artists <- list(names = c("Picasso", "Kahlo",
                genre = c("cubist", "surrealist"))
```

```{r}
# YOUR CORRECTED CODE HERE
```

#### 1.2 How many arguments does the `order()` function pass? What are they?

### 2. Vectors and Lists {-}

#### 2.1 Create three vectors: {-}

* a character vector, `titles`, that contain the names of 3 of your favorite movies

* a numeric vector, `year`, that contains the years in which those movies were produced

* a boolean vector `bechdel` that TRUE/FALSE according to whether those movies pass the [bechdel test](http://bechdeltest.com/)

```{r}
# YOUR CODE HERE
```

#### 2.2 Put those three vectors in a list, called `movies`. {-}

```{r}
# YOUR CODE HERE
```

#### 2.3 Print the structure of the list `movies` {-}

```{r}
# YOUR CODE HERE
```

### 3. Factors {-}

Here's some code that prints a simple barplot:

```{r}
f <- factor(c("low","medium","high","medium","high","medium"))
table(f)
barplot(table(f))
```

#### 3.1 How would you relevel `f` to be in the correct order? {-}

```{r}
# YOUR CODE HERE

# Test your code
barplot(table(f))
```

### 4. Dataframes {-}

#### 4.1 Coerce the `movies` object you made above from a list to a dataframe. Call it `movies_df` {-}

```{r}
# YOUR CODE HERE
```

#### 4.2 Add appropriate column names to `movies_df` {-}

```{r}
# YOUR CODE HERE
```