---
title: "Homework 3 Solutions"
author: "PLSC 31101"
date: "Fall 2019"
output: html_document
---

## Assignment 3

For this assignment, you'll be working on some real life data! I've prepared for your a basic country-year dataset, with the following variables:

* Country name
* Country numerical code
* Year
* UN Ideal point
* Polity2 score of regime type (from Polity VI)
* Physical Integrity Rights score (from CIRI dataset)
* Speech Rights score (from CIRI)
* GDP per capita (from World Bank)
* Population (from World Bank)
* Political Terror Scale using Amnesty International reports (from Political Terror Scale project)
* Composite Index of Military Capabilities (Correlates of War)
* Region

#### 1. R Projects and Importing {-}

__1.1 Initialize a project where you store this `.Rmd` file. Using `getwd()`, print your working directory below.__
```{r}
# YOUR CODE HERE
```

__1.2 Download the `country-year.csv` file and store it in a subdirectory of your project called `data`.__

__Read `country-year.csv` into R, using a relative path. Store it in an object called `dat`.__

```{r}
# YOUR CODE HERE
```

#### 2. Dimensions and Names {-}

__2.1 How many rows and columns are in the dataset?__

```{r}
# YOUR CODE HERE
```

__2.2 Print the column names.__

```{r}
# YOUR CODE HERE
```

__2.3 Remove the `X` column from the dataset.__

```{r}
# YOUR CODE HERE
```

__2.4 One of the variables is called "gdp.pc.wdi". This stands for "Gross Domestic Product Per Capita, from the World Bank Development Indicators". Change this variable name in the dataset from " "gdp.pc.wdi" to "GDP"__

```{r}
# YOUR CODE HERE
```

#### 3. Subsetting and Summarizing {-}

__3.1 Subset the data to include only observations from years 1990-2000 (inclusive). Put the subsetted data in a new variable called `dat.1990.2000`__

```{r}
# YOUR CODE HERE
```

__3.2 As a sanity check, summarize the `year` column of the subset you created above.__

```{r}
# YOUR CODE HERE
```

__3.3 Using `mean()` function, tell me the average GDP of observations from 1990 to 2000.

```{r}
# YOUR CODE HERE
```

__3.3 You just calculated the average GDP for years 1990-2000. Now calculate the average GDP from 2001 onwards. Tell me how much larger it is (in percentage).__

```{r}
# YOUR CODE HERE
```

__3.4  What does `summary(dat$polity2[dat$region =="Africa"])` do? Explain below in your own words.__



__3.5 Look up the helpfile for the function `is.na()`. Using this function, replace all the NA values in the `polity2` column of `dat` with 0.__ 

```{r}
# YOUR CODE HERE
```
