---
title: "Homework 2 Solutions"
author: "PLSC 31101"
date: "Fall 2019"
output: html_document
---

## Assignment 2 Solutions

- __Assigned__: Oct 10, 2019.
- __Due__: Oct 17, 2019 at 12:29pm.

For this assignment, you'll use what you know about R syntax and data structures to perform some common data operations.

#### 1. Basics {-}

__1.1 Fix the following syntax errors. Enter your corrected code in the second chunk.__

```{r eval = F}
# 1
states <- ("California", "Illinois", "Ohio")

# 2
countries <- c("Iran", "Indonesia," "India", "Italy")

# 3
df <- data.frame(age = c(21, 66, 35)
                 party = c('rep', 'dem', 'rep'))

# 4
my-vector <- c("apples", "oranges", "kiwis")

# 5
artists <- list(names = c("Picasso", "Kahlo",
                genre = c("cubist", "surrealist"))
```

```{r}
# PUT YOUR CORRECTED CODE HERE

# 1
states <- c("California", "Illinois", "Ohio")

# 2
countries <- c("Iran", "Indonesia", "India", "Italy")

# 3
df <- data.frame(age = c(21, 66, 35),
                 party = c('rep', 'dem', 'rep'))

# 4
my_vector <- c("apples", "oranges", "kiwis")

# 5
artists <- list(names = c("Picasso", "Kahlo"),
                genre = c("cubist", "surrealist"))
```

__1.2 How many arguments does the `order()` function pass? What are they?__

#### 2. Vectors and Lists {-}

__2.1 Create three vectors:__

* a character vector, `titles`, that contain the names of 3 of your favorite movies

* a numeric vector, `year`, that contains the years in which those movies were produced

* a boolean vector `bechdel` that TRUE/FALSE according to whether those movies pass the [bechdel test](http://bechdeltest.com/)

```{r}
titles <- c("Dog Day Afternoon", "The Graduate", "Breakfast Club")
year <- c(1975, 1967, 1985)
bechdel <- c(TRUE, FALSE, TRUE)
```

__2.2 Put those three vectors in a list, called `movies`.__

```{r}
movies <- list(titles, year, bechdel)
```

__2.3 Print the structure of the list `movies`.__

```{r}
str(movies)
```

#### 3. Factors {-}

__3.1 Here's some code that prints a simple barplot:__

```{r}
f <- factor(c("low","medium","high","medium","high","medium"))
table(f)
barplot(table(f))
```

__How would you relevel `f` to be in the correct order?__

```{r}
f <- factor(f, levels = c("low", "medium", "high"))

# Test your code
barplot(table(f))
```

#### 4. Dataframes {-}

__4.1 Coerce the `movies` object you made above from a list to a dataframe. Call it `movies_df`.__

```{r}
movies_df <- as.data.frame(movies)
```

__4.2 Add appropriate column names to `movies_df`.__

```{r}
names(movies) <- c("film", "year", "bechtel")
```