---
title: "Homework 3 Solutions"
author: "PLSC 31101"
date: "Fall 2019"
output: html_document
---

## Assignment 3 Solutions

- __Assigned__: Oct 17, 2019.
- __Due__: Oct 24, 2019 at 12:29pm.

For this assignment, you'll be working on some real life data! I've prepared for your a basic country-year dataset, with the following variables:

* Country name
* Country numerical code
* Year
* UN Ideal point
* Polity2 score of regime type (from Polity VI)
* Physical Integrity Rights score (from CIRI dataset)
* Speech Rights score (from CIRI)
* GDP per capita (from World Bank)
* Population (from World Bank)
* Political Terror Scale using Amnesty International reports (from Political Terror Scale project)
* Composite Index of Military Capabilities (Correlates of War)
* Region

#### 1. R Projects and Importing {-}

__1.1 Using `getwd()`, print your working directory below.__
```{r}
getwd()
```

__1.2 Read `country-year.csv` into R, using a relative path. Store it in an object called `dat`.__

```{r}
dat <- read.csv("data/country-year.csv")
```

#### 2. Dimensions and Names {-}

__2.1 How many rows and columns are in the dataset?__

```{r}
dim(dat)
```

__2.2 Print the column names.__

```{r}
names(dat)
```

__2.3 Remove the `X` column from the dataset.__

```{r}
dat$X <- NULL
```

__2.4 One of the variables is called "gdp.pc.wdi". This stands for "Gross Domestic Product Per Capita, from the World Bank Development Indicators". Change this variable name in the dataset from " "gdp.pc.wdi" to "GDP"__

```{r}
names(dat)[8] <- "GDP"
```

#### 3. Summarizing {-}

__3.1 How many years are covered in the dataset?__

```{r}
length(unique(dat$year))
```

__3.2 How many unique countries are covered in the dataset?__

```{r}
length(unique(dat$country))
```

__3.3 What is the range of `polity2`? How many NAs are in this column?__

```{r}
summary(dat$polity2)
```


#### 4. Subsetting {-}

__4.1 Subset dat so that it returns the third column AS A VECTOR__ (Do not print the object; store it in a variable.)

```{r}
sub <- dat[[3]]
#OR
sub <- dat[,3]
#OR
names(dat)[3]
sub <- dat$country
```

__4.2 Fix each of the following common data frame subsetting errors__:

1. Extract observations collected for the year 1980

```{r, eval=FALSE}
dat[dat$year = 1980,]

# Corrected
dat[dat$year == 1980,]
```

2. Extract all columns except 1 through to 4

```{r, eval=FALSE}
dat[,-1:4]

# Corrected
dat[,-c(1:4)]
```

3. Extract the rows where the polity2 score is greater than 5

```{r, eval=FALSE}
dat[dat$polity2 > 5]

# corrected
dat[dat$polity2 > 5, ]
```

4. Extract the first row, and the third and fourth columns
  (`country` and `idealpoint`).

```{r, eval=FALSE}
dat[1, 3, 4]

# Corrected
dat[1, c(3, 4)]
```

5. Extract rows that contain information for the years 2002
   and 2007

```{r, eval=FALSE}
dat[dat$year == 2002 | 2007,]

# Corrected
dat[dat$year == 2002 | dat$year == 2007,]
```

__4.3  What does `summary(dat$polity2[dat$region =="Africa"])` do? Explain below in your own words.__

It calculates some summary statistics for `polity2` scores from observations in Africa.

__4.4 Subset the data to include only observations from years 1990-2000 (inclusive). Put the subsetted data in a new variable called `dat.1990.2000`__

```{r}
dat.1990.2000 <- dat[dat$year >= 1990 & dat$year<=2000,]
```

__4.5 Using `mean()` function, tell me the average GDP of observations from 1990 to 2000.__

```{r}
mean(dat.1990.2000$GDP, na.rm = T)
```

__4.6 You just calculated the average GDP for years 1990-2000. Now calculate the average GDP from 2001 onwards. Tell me how much larger it is (in percentage).__

```{r}
dat.2001.plus <- dat[dat$year > 2000,]
mean1 <- mean(dat.1990.2000$GDP, na.rm = T)
mean2 <- mean(dat.2001.plus$GDP, na.rm = T)
(mean2 - mean1) / mean1
```

__4.7 Look up the helpfile for the function `is.na()`. Using this function, replace all the NA values in the `polity2` column of `dat` with 0.__ 

```{r}
?is.na
dat$polity2[is.na(dat$polity2)] <- 0
summary(dat$polity2)
```
