---
title: "Homework 4"
author: "PLSC 31101"
date: "Fall 2019"
output: html_document
---

## Assignment 4 Solutions

- __Assigned__: Oct 24, 2019
- __Due__: Nov 5, 2019 at 12:29pm.

For this problem set, we'll be working with the country-year data introduced in the last assignment. As a reminder, the dataset contains the following variable:

* `year`: Year.
* `ccode`: Country numerical code.
* `country`: Country name.
* `idealpoint` UN Ideal point.
* `polity2`: Polity2 score of regime type (from Polity VI).
* `physint`: Physical Integrity Rights score (from CIRI dataset).
* `speech`: Speech Rights score (from CIRI).
* `gdp.pc.wdi`: GDP per capita (from World Bank).
* `pop.wdi`: Population (from World Bank).
* `amnesty`: Political Terror Scale using Amnesty International reports (from Political Terror Scale project).
* `cinc`: Composite Index of Military Capabilities (Correlates of War).
* `region`: Geographic region.

We'll be merging this country_year data with new data about U.S. news coverage of women around the world (excluding the United States). In this new dataset, the unit of observation is *article*. That is, each row represents an individual article, with columns for:

* `publication`: NYT or Washington Post.
* `year`: Year article was published.
* `title`: Title of the article.
* `country`: Country the article is mainly about.
* `region`: Region where `country` is located.
* `ccode`: Numerical code for `country`.

#### 1. Loading, subsetting, summarizing {-}

__1.1 Load the csv found in `data/articles.csv` into R. Be sure to set `stringsAsFactors` to `FALSE`. Store the data-frame to an object called `articles` and tell me the variable names.__

```{r}
library(tidyverse)
articles <- read.csv("data/articles.csv", stringsAsFactors = F)
names(articles)
```

__1.2 How many countries are covered in the dataset?__

```{r}
length(unique(articles$country))
```

__1.3 The variable `ccode` reports a numerical ID corresponding to a given country. Print the names of the country or countries without a `ccode` (i.e. those countries where the `ccode` is `NA`.)__

```{r}
unique(articles$country[is.na(articles$ccode)])
```

__1.4 Remove all articles where the `ccode` variable is `NA`. How many observations are left with?__

```{r}
articles_no_na <- articles[!is.na(articles$ccode), ]
nrow(articles_no_na)
```

#### 2. Counting Frequencies and Merging {-}

__2.1 Create a new data frame called `articles_country_year` that tells us the number of articles per `ccode` (i.e. country code), per `year`.__

__The final data frame `articles_country_year` should contain three columns: `ccode`, `year`, and `number_articles`.__

__Print the first 6 rows of the `articles_country_year`.__

**Hint**: The `count` function -- part of the `plyr` package -- might be helpful.

```{r}
articles_country_year <- articles_no_na %>% 
  dplyr::count(ccode, year) %>%
  select(ccode, year, number_articles = n)

kable(head(articles_country_year))
```

__2.2. Load `data/country-year.csv` (this is the country-year data we worked with during the last assignment.)__

```{r}
country_year <- read.csv("data/country-year.csv", stringsAsFactors = F)
```

__2.3  Subset `country_year` such that it has the same year range as `articles_country_year`.__

```{r}
range(articles_country_year$year)
range(country_year$year)

country_year <-country_year %>% 
  filter(year > 1979)
```

__2.4  Merge (i.e. join) `articles_country_year` and `country_year` into a new dataframe called `merged`.__ 

__When you're done, `merged` should have all the rows and columns of the country_year dataset, along with a new column called `number_articles`.__

__Print the first 6 rows of this new dataframe `merged`.__

```{r}
merged <- country_year %>% 
  left_join(articles_country_year)

kable(head(merged))
```

__2.5 In `merged`, replace all instances of `NA` in the `number_articles` column to 0.__

```{r}
# solution 1 - base R
merged$number_articles[is.na(merged$number_articles)] <- 0

# solution 2 - tidyr
merged$number_articles <- replace_na(merged$number_articles, 0)

# solution 3 - dplyr
merged <- merged %>% 
  mutate(number_articles = ifelse(is.na(number_articles), 0, number_articles))

# test
summary(merged$number_articles)
```

__2.6  Which country-year observation has the most number of articles? Write code that prints the year, country name, and number of articles for this observation.__

```{r}
# solution #1 -- base R
merged[which.max(merged$number_articles),c("year", "country", "number_articles")]

# solution #2 -- tidyverse
merged %>% 
  top_n(1, number_articles) %>%
  select(year, country, number_articles)
```


#### 3. Group-wise Operations {-}

__3.1 Using the `merged` data and our **split-apply-combine** strategies, print the total number of articles about women per region.__

```{r}
n_region <- merged %>%
  group_by(region) %>%
  summarise(count = sum(number_articles, na.rm = T))

n_region
```

#### 4. Long v. wide formats {-}

__4.1 Create a piped operation on `merged` that does the following:__

  __1. Subsets the dataframe to select `year`, `country`, and `number_articles` columns.__
  __2. Filters the dataframe to select only observations in the MENA region.__
  __3. Spreads the dataframe so that each `country` is a column, and the cells represent `number_articles.__

__Print the first 6 rows of this transformed data frame.__

```{r}
wide <- merged %>%
  filter(region == "MENA") %>%
  select(year, country, number_articles) %>%
  spread(country, number_articles, fill = 0)

kable(head(wide))
```

__4.2 Transform the dataset you created above back into `long` format, with three variables: `year`, `country`, and `number_articles`__

__Print the first 6 rows of this transformed data frame.__

```{r}
long <- wide %>% 
  gather(country, number_articles, -year)

kable(head(long))
```

#### Extra Credit

__This question is not required. But it you want an extra challenge....__

__Transform the `country_year` data into an undirected dyadic dataset. Here, the unit of observation should be the dyad-year, with five columns:__

  1. `ccode_1`: Country 1 `ccode`
  2. `country_1`: Country 1 name
  3. `ccode_2`: Country 2 `ccode`
  4. `country_2`: Country 2 name
  5. `year`: Year of observation
  3. `gdp_diff`: Absolute difference of gdp between dyad. 

__This should be *undirected* dyadic dataset, meaning `USA-Canada-1980` is the same as `Canada-USA-1980`, and we shouldn't have duplicate rows for the same dyad.__

__Try to do it all in 1 piped sequence. Then tell me the dyad-year with the greatest wealth disparity.__

```{r}
dyad <- country_year %>% 
  expand(ccode_1=ccode, ccode_2=ccode) %>% # make two columns of states
  filter(ccode_1 > ccode_2) %>% # from directed to undirected dyads
  left_join(., country_year, by=c("ccode_1"="ccode")) %>% # get state1 info
  left_join(., country_year, by=c("year", "ccode_2"="ccode")) %>% # get state2 info 
  mutate(gdp_diff = abs(gdp.pc.wdi.x - gdp.pc.wdi.y)) %>% # take absolute difference in gdp
  select(ccode_1, country_1 = country.x, ccode_2, country_2 = country.y, year, gdp_diff) %>%
  arrange(desc(gdp_diff))

kable(head(dyad))
```