---
title: "Basic Syntax"
author: "PLSC 31101"
date: "Fall 2019"
output: html_document
---

## Data Types

Every value in a program has a specific __type__. In R, those types are called "classes", and there are 4 of them:

* character (text or "string")
* numeric (integer or decimal)
* integer (just integer)
* logical (TRUE or FALSE booleans)

| Example | Type |
| ------- | ---- |
| "a", "swc" | character |
| 2, 15.5 | numeric | 
| 2 (Must add a `L` at end to denote integer) | integer |
| `TRUE`, `FALSE` | logical |

### What's that Type?

R is dynamically typed, meaning that it "guesses" what class a value is. 

Use the built-in function```class()``` to find out what type a value has.

```{r}
class(3)
class(3L)
class("Three")
class(T)
```

This works on variables as well. But remember: the _value_ has the type --- the _variable_ is just a label.

```{r}
three <- 3
class(three)

three <- "three"
class(three)
```

A value's class determines what the program can do to it.

```{r eval = F}
3 - 1
3 - "1"
```

### Coercion

We just learned we cannot subtract numbers and strings. Instead, use  `as..` + name of class as functions to convert a value to that type.

```{r eval = F}
3 - as.numeric("1")
```

This is called `coercion`. Here's another example:

```{r}
my_var <- "FALSE"
my_var
as.logical(my_var)
```

What difference did you notice?

### Other Objects

There are a few other "odd ball" types in R:

#### `NA` are missing values {-}

Missing values are specified with `NA`. `NA` will always be coerced to the correct type if used inside `c()`

```{r}
x <- c(NA, 1)
x
typeof(NA)
typeof(x)
```

#### `Inf` is infinity. {-}

You can have either positive or negative infinity.

```{r, eval = TRUE}
1/0
1/Inf
```

#### `NaN` means "Not a number". It's an undefined value. {-}

```{r, eval = TRUE}
0/0
```


### Challenges

#### Challenge 1: Making and Coercing Variables {-}

1. Make a variable `year` and assign it as the year you were born.
2. Coerce that variable to a string, and assign it to a new variable `year_string`.
3. Someone in your class says they were born in 2001. Really? Really. Find out what your age difference is, using only `year_string`

```{r}
year <- 1992
year_string <- as.character(year)
2001 - as.numeric(year_string)
```


#### Challenge 2: Fix the Code {-}

Change the following code to make the output TRUE.

```{r}
val_1 = F
val_2 = "F"

class(val_1) == class(val_2)
```

Solution:
```{r}
val_1 = F
val_2 = "F"

class(as.character(val_1)) == class(val_2)
```

